BOLD Parameter Estimation Using Particle Filters

### 2010-08-16 doc/thesis/methods.tex
spelling fixes etc.

\section{Model}
As originally written in \autoref{sec:BOLD Physiology} the state variables
for the BOLD model are as follows:
\begin{eqnarray}
\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
\dot{f} &=& s\\
\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
\end{eqnarray}
The original assumption regarding particle filter models (\autoref{sec:Particle Filter Model})
included noise in the update of $x$, however that is not included here.
The reason for the difference is that cloud of particles is, to some extent,
able to account for that noise. It is common, however, to model that noise
in particle filters by adding a random value to each updated state variable. 
Because the purpose of this particle filter is to learn the underlying distribution
of the static parameters, rather than precisely modeling the time course of the 
in the dynamic state variables $\{s,f,v,q\}$ this noise is left out. It also helps
that detrending is applied before the particle filter and that the
BOLD model is dissipative. When no stimuli are applied, all the particles 
decay to ($\{0,1,1,1\}$). Typical particle filters 
also use this state noise as an exploratory measure; however this method is
less necessary when good priors are available.

For all the analyses  in this work, $1400$ integration points
per second were used.  Typically a step size of $0.001$ was sufficient,
however, from time to time $0.001$ can still result in problems for the
BOLD model.

\section{Preprocessing}
\label{sec:Methods Preprocessing}
The normal pipeline for analyzing
FMRI involves a several preprocessing steps. The first and most important
task is motion correction. To do this, a single volume in time is chosen, and
volumes at every other time are realigned to best match the target volume. This corrects
for motion by the patient as well as small changes in the magnetic
fields that cause the image to shift. 
In conventional statistical parametric mapping, a Gaussian smoothing
filter is applied across the image as discussed in \autoref{sec:RFT}.
After this, detrending is performed which is discussed in \autoref{sec:Detrend}.
Recall that FMRI signal levels are unit-less and though detrending is not
always necessary, the data must always be converted 
into \% difference from baseline. 
The generally accepted method is to use a high pass filter, although the
cutoff frequency is application dependent and often applied haphazardly.
Before going into the detrending used in this work, it is necessary to 
discuss the type of noise present in FMRI.

\subsection{BOLD Noise}
\label{sec:Introduction Noise}
As demonstrated in \autoref{sec:BOLD Physiology} the BOLD response has been
extensively studied and despite minor discrepancies, the cause of the BOLD 
signal is well known. However, as FMRI detects an  
aggregate signal over the space of cubic centimeters, there are
plenty of noise sources . Though local neurons act
together (i.e. around the same time), the density of neurons, the
density of capillaries, and slight differences in activation across 
a particular voxel can all lead to signal attenuation and noise.

@@

In order to characterize the noise, I analyzed resting state data.
During resting state, the patient is shown no images, and he is asked
to avoid movement and complex thought.  Overall though there should be 
very little activation, and thus the signal consists entirely of noise. 
Therefore resting state data is perfect for analyzing noise. 
The locations were chosen from points all around the brain, 
all in grey matter voxels. These time
series were chosen because they were representative of different types
of noise found in the resting state data.

@@

Because most methods (including the one used in this paper)
assume the noise realizations are independent of each other, the autocorrelation 
is of particular interest (which is a necessary but not
sufficient condition for independence). Gaussianity is also a common
assumption made in studies of FMRI data, though that assumption is not
needed in this work. Regardless, comparing the distribution to a Gaussian
is informative, so Q-Q plots are used to compare example data with the
Normal distribution. Additionally, in FMRI data the noise is often considered 
to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
characterized by steps that are Gaussian and independent. The simulations discussed in 
\autoref{sec:Single Voxel Simulation} make use of this, 
by adding a Wiener random process to the overall signal. To determine
whether the noise is in fact Wiener, the distribution of 
the steps were also plotted against a Gaussian.

@@

De-trending the time-series by subtracting a spline fit to the distribution
removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
though not perfectly. Though the distributions still do not exactly fit
the Normal, \autoref{fig:QQs:D} is much improved compared to \autoref{fig:QQDC:D}.
In all, the de-trending is effectively removing Wiener noise.

\subsection{Detrending}
\label{sec:Detrend}
The non-stationary
aspect of a Weiner process, presumably the result of integrating some
$\nu_x$ is difficult to compensate for, and so many methods
have been developed to compensate for it. Tanabe et al. and Smith et al. 
demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
of FMRI \cite{Tanabe2002, Smith1999}. It has been reported that in some studies as 
many as half the voxels 
benefited from detrending \cite{Smith2007}. In a head to head comparison, 
Tanabe et al.  showed that in most cases subtracting off
a spline worked the best \cite{Tanabe2002}. 
The benefit of the spline versus wavelets, high pass 
filtering or other DC removal techniques is that the frequency response is not set.
Rather, the spline is adaptive to the input. Unfortunately no method will 
perfectly remove noise, and no method will leave the signal untouched.

The method I used to calculate the spline was picking one knot for every 20
measurements in an image. Thus a 10 minute session at a repetition time of 
2.1 seconds would have 19 knots. The first and last knots were each 
given half the number of samples as the rest of the knots; which were all 
located at the center of their sample group. The median of each sample group
was then taken and used as the magnitude for the group. Taking the median 
versus the mean seemed to work better, given the presence of outliers. 
There is potential to optimize the spline further using a canonical 
HRF to find resting points; however, the experiment would have
to be designed with this in mind.

Problematically, after removing the DC component of the signal,
by definition the signal will have a median near zero. 
Unfortunately this is not the natural state of the BOLD signal. More specifically,
when the signal is inactive, the BOLD response should be at 0\% change from
the base level; activation may then increase, or for short periods decrease from this base.
Because most of the BOLD signal is above baseline, after removing the spline
the resting state will be below 0\%.  
One method of accounting for this is to simply add a DC gain model parameter.
Like all the other model parameters, with enough measurements, the particle filter
would be able to settle on a good estimate. Yet adding another dimension increases the
complexity of the model, when the parameter is relatively easy to estimate
by visual inspection.  In this work a simpler approach was used. To determine
the DC gain I used a robust estimator of scale. The Median Absolute Deviation (MAD)
proved to be accurate in determining how much to shift the signal up
by. I tested both methods during the course of analysis, and found that the increase 
in model complexity far outweighed the slight increase in flexibility. Other
methods may work better, however the MAD worked well, 
as \autoref{fig:PreprocessedLowNoise} and \autoref{fig:PreprocessedHighNoise} show.

@@

%\subsection{Linearizing Noise}
%\label{sec:Methods Delta Based Inference}
%The alternative to these sorts of low frequency manipulation is to
%go around the problem in another way. Here, I propose a 
%different method of dealing with the drift. 
%Instead of comparing the direct output of the particle filter with the direct
%measurement, the algorithm would compare the change in signal over a single TR,
%with the result of integrating the model for the same period. 
%In most signal processing cases this would be foolish, but that is because the 
%general assumption is that all noise is high frequency. Considering 
%the fact that every BOLD analysis pipeline uses a high pass filter,
%whereas low poss temporal filter are rarely applied, it makes sense
%that a derivative type method could work. The benefit of particle filters
%is that they are a robust method of inference, and I would assert 
%that the particle filter ought to be given as \emph{raw} data as possible. 
%While taking direct measurements
%without de-trending would give awful results, using the difference removes the 
%DC component and turns what is usually assumed to be a Weiner process into 
%a simple Gaussian random variable. 
%
%\begin{equation}
%\Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
%\label{eq:measass_delta}
%\end{equation}
%
%Even if $\nu_d$ is some other additive process, the difference will still be closer
%to I.I.D. than a Wiener process, as the autocorrelation of the $\delta y$ shows
%in \autoref{fig:QQDelta} in \autoref{sec:Introduction Noise}. 
% All the assumptions made originally
%for the particle filter still hold, and all of the parameters may be distinguished based on
%the step sizes, thus it is not unreasonable to consider matching the string of step sizes
%rather than string of direct readings. 
%
%\begin{figure}
%\label{fig:FrequencyGraphs}
%\caption{frequency response graphs, highlighting noise frequency range and signal frequency range}
%\end{figure}

@@

\subsection{Prior Distribution}
\label{sec:PriorDist}
For the BOLD model described in \autoref{sec:BOLD Physiology}, several
different studies have endeavored to calculate parameters. The results
of these studies may be found in \autoref{tab:Params}, and the methods 
used for each may be found in \autoref{sec:Prior Works}. Unfortunately,
Friston et al. only studied regions deemed active by the General 
Linear Model; and most other research (including \cite{Friston2001}) used these results as 
the source for their priors \cite{Friston2000}. 
The one exception is to this is Johnston et al. which came to a extremely different
distributions \cite{Johnston2008}. For a particle filter, the choice of a prior is
the single most important design decision. A very wide prior will require
more particles to be sufficiently dense, a very thin (low variance) prior may miss
the true parameters. Consequently, for this work it was natural
to use priors that would give results consistent with previous works
\cite{Friston2000}. This constrains the usefulness of the model to
areas that fall within the prior distribution, yet allowed results
to be comparable to other works. There is a significant need for better
estimates of the physiological parameters; and, while physical experiments
may not be possible, it would not be unreasonable to do a study with
exhaustive simulated annealing or hill climbing tests for multiple
regions and multiple patients. That said, a significant amount of
indeterminability exists in the parameters, which is discussed in 
\autoref{sec:Single Voxel Simulation}

@@

It is also import to use enough particles to obtain a 
sufficiently dense approximation of the prior. For 7 dimensions,
getting a dense prior is difficult. Insufficiently
dense particles will result in inconsistent results. Of course the
processing time will scale up directly with the number of particles.
A dense initial estimate is important so that some particles land
near the solution; but as the variance decreases the number of 
particles needed decreases as well. Thus, as a heuristic, initially
the number of particles was set to 16,000, but after resampling,
the number of particles was dropped to 1,000. Typically during the 
first few measurements the variance drops precipitously because
most particles are far from a solution.  The particles that are left are in a
much more compact location, allowing them to be estimated with 
significantly fewer particles. These numbers aren't set in stone,
and depending on the complexity of the system or desired accuracy
they could be changed; however, they seem to be the minimum that
will give consistent results.

@@

Resampling is a not strictly necessary, but it increases the effectiveness
of the particle filter by adjusting the support to emphasize areas
of higher probability. Resampling is slow because it requires redrawing
all the particles. It also closes off avenues of investigation, and is
designed to over-smooth to prevent overly thinning the support. For all these
reasons, resampling was only performed when the $N_{eff}$ dropped below
50 (for 1000 particles).  As a measure against sharp drops in the $N_{eff}$ 
caused by a large spike in error, resampling was only performed when 
two consecutive low ($<50$) $N_{eff}$'s were found.

\subsection{Particle Deprivation}
\label{sec:Particle Deprivation}
To prevent particle deprivation, which may happen even with a good
estimate of the prior distribution, a method of rescuing the particle
filter from this state was used. When particle deprivation was detected,
resampling was performed with a saved version of the covariance matrix (from
the previous time step). This allowed for the particles to be re-scattered
without having to go all the way back to the prior distribution. Particle 
deprivation was defined by $N_{eff}$ dropping to $1$.

@@

After trial and error, for this work I chose the Gaussian with a standard deviation 
of $0.005$. Although using an adaptive weighting function
might be better, in tests this often led to unpredictable results. 
With some work it may be possible to set the standard deviation by
taking a small sample from resting data and using the sample standard deviation.
In this work, however, I set the standard deviation manually at $0.005$,
because it gave the best consistency.

\subsection{Runtime}
The run-time for a single voxel depends on the several factors. First, the
overall length of the signal being analyzed. For 1000 measurements it takes
approximately 6 minutes. On the other hand, in real circumstances the
length tends to be around 150 measurements which takes around 40 seconds (for 1000 
particles, 1400 integration points and a Quad Core CPU). The size of 
integration steps are also crucial, but using step sizes above 
$0.001$ seconds is not recommended. In most cases millisecond resolution
is fine; however, when generating simulated data I found that at times it was
still not enough. This is problematic in the actual particle
filter since, given the large number of simultaneous integrations taking 
place, its probable that a few particles will fail and be unfairly thrown away.
To prevent such events, 1400 integration points per second were used throughout
the tests.

>>>>

\section{Model}
As originally written in \autoref{sec:BOLD Physiology} the state variables
for the BOLD model are as follows:
\begin{eqnarray}
\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
\dot{f} &=& s\\
\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
\end{eqnarray}
The original assumption regarding particle filter models (\autoref{sec:Particle Filter Model})
included noise in the update of $x$, however that is not included here.
The reason for the difference is that the cloud of particles is, to some extent,
able to account for that noise. It is common, however, to model that noise
in particle filters by adding a random value to each updated state variable.
Because the purpose of this particle filter is to learn the underlying distribution
of the static parameters, rather than precisely modeling the time course of the
in the dynamic state variables $\{s,f,v,q\}$ this noise is left out. It also helps
that detrending is applied before the particle filter and that the
BOLD model is dissipative. When no stimuli are applied, all the particles
decay to $\{0,1,1,1\}$. Typical particle filters
also use this state noise as an exploratory measure; however, this method is
less necessary when good priors are available, and when regularized particle filtering
is being used.

For all the analysis  in this work, $1400$ integration points
per second were used.  Typically a step size of $0.001$ was sufficient,
however, from time to time $0.001$ can still result in problems for the
BOLD model.

\section{Preprocessing}
\label{sec:Methods Preprocessing}
The normal pipeline for analyzing
FMRI involves a several preprocessing steps. The first and most important
task is motion correction. To do this, a single volume in time is chosen, and
volumes at every other time point are realigned to best match the target volume. This corrects
for motion by the patient as well as small changes in the magnetic
fields that cause the image to shift.
In conventional statistical parametric mapping, a Gaussian smoothing
filter is applied across the image as discussed in \autoref{sec:RFT}.
After this detrending is performed, which is discussed in \autoref{sec:Detrend}.
Recall that FMRI signal levels are unit-less and though detrending is not
always necessary, the data must always be converted
into \% difference from baseline.
The generally accepted method is to use a high pass filter, although the
cutoff frequency is application dependent and often applied haphazardly.
Before going into the detrending used in this work, it is necessary to
discuss the type of noise present in FMRI.

\subsection{BOLD Noise}
\label{sec:Introduction Noise}
As demonstrated in \autoref{sec:BOLD Physiology} the BOLD response has been
extensively studied and despite minor discrepancies, the cause of the BOLD
signal is well known. However, as FMRI detects an
aggregate signal over the space of cubic centimeters, there are
plenty of noise sources. Though local neurons act
together (i.e. around the same time), the density of neurons, the
density of capillaries, and slight differences in activation across
a particular voxel can all lead to signal attenuation and noise.

@@

In order to characterize the noise, I analyzed resting state data.
During resting state, the patient is shown no images, and he is asked
to avoid movement and complex thought.  Overall there should be
very little activation, and thus the signal consists entirely of noise.
Therefore resting state data is perfect for analyzing noise.
The locations were chosen from points all around the brain,
all in grey matter voxels. These time
series were chosen because they were representative of different types
of noise found in the resting state data.

@@

Because most methods (including the one used in this paper)
assume the noise realizations are independent of each other, the autocorrelation
is of particular interest (which is a necessary but not
sufficient condition for independence). Gaussianity is also a common
assumption made in studies of FMRI data, though that assumption is not
needed in this work. Regardless, comparing the distribution to a Gaussian
is informative, so Q-Q plots are used to compare example data with the
Normal distribution. Additionally, in FMRI data, the noise is often considered
to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
characterized by steps that are independent and Gaussian. The simulations discussed in
\autoref{sec:Single Voxel Simulation} make use of this,
by adding a Wiener random process to the overall signal. To determine
whether the noise is in fact Wiener, the distribution of
the steps were also plotted against a Gaussian.

@@

Detrending the time-series by subtracting a spline fit to the distribution
removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
though not perfectly. Though the distributions still do not exactly fit
the Normal, \autoref{fig:QQs:D} is much improved compared to \autoref{fig:QQDC:D}.
In all, the detrending is effectively removing Wiener noise.

\subsection{Detrending}
\label{sec:Detrend}
The non-stationary
aspect of a Weiner process, presumably the result of integrating some
$\nu_x$, is difficult to compensate for, and so many methods
have been developed to compensate for it. Tanabe et al. and Smith et al.
demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
of FMRI \cite{Tanabe2002, Smith1999}. It has been reported that in some studies as
many as half the voxels
benefited from detrending \cite{Smith2007}. In a head to head comparison,
Tanabe et al.  showed that in most cases subtracting off
a spline worked the best \cite{Tanabe2002}.
Unfortunately no method will
perfectly remove noise, and no method will leave the signal untouched.

The method I used to calculate the spline was one knot for every 20
measurements in an image. Thus a 10 minute session at a repetition time of
2.1 seconds would have 19 knots. The first and last knots were each
given half the number of samples as the rest of the knots; which were all
located at the center of their sample group. The median of each sample group
was then taken and used as the magnitude for the group. Taking the median
versus the mean seemed to work better, given the presence of outliers.
There is potential to optimize the spline further using a canonical
HRF to find resting points; however, the experiment would have
to be designed with this in mind.

Problematically, after removing the DC component of the signal,
by definition the signal will have a median near zero.
Unfortunately this is not the natural state of the BOLD signal. More specifically,
when the signal is inactive, the BOLD response should be at 0\% change from
the base level; activation may then increase, or for short periods decrease from this base.
Because most of the BOLD signal is above baseline, after removing the spline
the resting state will be below 0\%.
One method of accounting for this is to simply add a DC gain model parameter.
Like all the other model parameters, with enough measurements, the particle filter
would be able to settle on a good estimate. Yet adding another dimension increases the
complexity of the model, for a parameter that is relatively easy to estimate
by visual inspection.  In this work a simpler approach was used. To determine
the DC gain I used a robust estimator of scale. The Median Absolute Deviation (MAD)
proved to be accurate in determining how much to shift the signal up
by. I tested both methods during the course of analysis, and found that the increase
in model complexity far outweighed the slight increase in flexibility. Other
methods may work better, however the MAD worked well,
as \autoref{fig:PreprocessedLowNoise} and \autoref{fig:PreprocessedHighNoise} show.

@@

@@

\subsection{Prior Distribution}
\label{sec:PriorDist}
For the BOLD model described in \autoref{sec:BOLD Physiology}, several
different studies have endeavored to calculate parameters. The results
of these studies may be found in \autoref{tab:Params}, and the methods
used for each may be found in \autoref{sec:Prior Works}. Unfortunately,
Friston et al. only studied regions deemed active by the General
Linear Model; and most other research used these results as
the source for their priors \cite{Friston2002}.
The one exception is to this is Johnston et al. which came to a extremely different
distributions \cite{Johnston2008}. For a particle filter, the choice of a prior is
the single most important design decision. A very wide prior will require
more particles to be sufficiently dense, a very thin (low variance) prior may miss
the true parameters. Consequently, for this work it was natural
to use priors that would give results consistent with previous works
\cite{Friston2002}. This constrains the usefulness of the model to
areas that fall within the prior distribution, yet allowed results
to be comparable to other works. There is a significant need for better
estimates of the physiological parameters; and, while physical experiments
may not be possible, it would not be unreasonable to do a study with
exhaustive simulated annealing or hill climbing tests for multiple
regions and multiple patients. That said, a significant amount of
indeterminateness exists in the parameters, which is discussed in
\autoref{sec:Single Voxel Simulation}

@@

It is also import to use enough particles to obtain a
sufficiently dense approximation of the prior. For 7 dimensions,
getting a dense prior is difficult. Insufficiently
dense particles will result in inconsistent results. Of course the
processing time will scale up directly with the number of particles.
A dense initial estimate is important to ensure that some particles land
near the solution; but as the variance decreases the number of
particles needed decreases as well. Thus, as a heuristic, initially
the number of particles was set to 16,000, but after resampling,
the number of particles was dropped to 1,000. Typically during the
first few measurements the variance drops precipitously because
most particles are far from a solution.  The particles that are left are in a
much more compact location, allowing them to be estimated with
significantly fewer particles. These numbers aren't set in stone,
and depending on the complexity of the system or desired accuracy
they could be changed; however, they seem to be the minimum that
will give consistent results.

@@

Resampling is a not strictly necessary, but it increases the effectiveness
of the particle filter by adjusting the support to emphasize areas
of higher probability. Resampling is slow because it requires redrawing
all the particles. It also closes off avenues of investigation, but is
designed to over-smooth to prevent overly thinning the support. For all these
reasons, resampling was only performed when the $N_{eff}$ dropped below
50 (for 1000 particles).  As a measure against sharp drops in the $N_{eff}$
caused by a large spike in error, resampling was only performed when
two consecutive low ($<50$) $N_{eff}$'s were found.

\subsection{Particle Deprivation}
\label{sec:Particle Deprivation}
To prevent particle deprivation (all weights dropping to 0), which may happen even with a good
estimate of the prior distribution, a method of rescuing the particle
filter from this state was implemented. When particle deprivation was detected,
resampling was performed with a saved version of the covariance matrix (from
the previous time step). This allowed for the particles to be re-scattered
without having to go all the way back to the prior distribution. Particle
deprivation was defined by $N_{eff}$ dropping to $1$.

@@

After trial and error, for this thesis I chose the Gaussian with a standard deviation
of $0.005$. Although using an adaptive weighting function
might be better, in tests this often led to unpredictable results.
With some work it may be possible to set the standard deviation by
taking a small sample from resting data and using the sample standard deviation.
In this work, however, I set the standard deviation manually at $0.005$,
because it gave the best consistency.

\subsection{Runtime}
The run-time for a single voxel depends on the several factors. First, the
overall length of the signal being analyzed. For 1000 measurements it takes
approximately 6 minutes. On the other hand, in real circumstances the
length tends to be around 150 measurements which takes around 40 seconds (for 1000
particles, 1400 integration points and a Quad Core CPU). The size of
integration steps are also crucial. Using step sizes above
$0.001$ seconds is not recommended. In most cases millisecond resolution
is fine; however, when generating simulated data I found that at times it was
still not enough. This is problematic in the actual particle
filter since, given the large number of simultaneous integrations taking
place, its probable that a few particles will fail and be unfairly thrown away.
To prevent such events, 1400 integration points per second were used throughout
the tests.

### 2010-08-16 doc/thesis/results.tex
FINAL!
spelling fixes etc.

To test how well the particle filter would do with plenty of data, and 
determine the inherent variance in the model parameters, a very long simulation
with randomly generated impulse stimuli was created. The preprocessed timeseries
and the final estimate are shown in \autoref{fig:long_converge};
the final covariance matrix of the parameters is in \autoref{tab:long_cov}.
Note that even though the BOLD response converged (\autoref{fig:long_converge},
\autoref{fig:long_converge_500}),
the parameters still have significant correlation (\autoref{tab:long_corr}). 
Based on the histograms in the first 500 seconds, the parameters converged to
their final values well before the end.
Although this is only a single test, the correlation (\autoref{tab:long_corr}) 
of the parameters 
indicates significant parameter indeterminability. When the input
consists entirely of impulses, the best parameters are not one particular
set, but a joint distribution. Note that the correlation is in parameters whose priors
are completely independent. It is possible that varying the type of input could
give improved results, although informal tests did not show significant difference.
In spite of the noisy input (green line in \autoref{fig:long_converge}), the estimates
of the BOLD were actually very close to the true (noise-free) BOLD signal.

@@

Note that throughout the results, unless otherwise specified,
the term \emph{residual} will be defined as the square root of the mean squared residual, 
\begin{equation}
\text{Residual} = \sqrt{\sum (\hat{y}_k - y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the preprocessed output
sampled at time $k$. Note the subtle distinction between this and what will be 
termed the \emph{error} which is defines as 
\begin{equation}
\text{Error} = \sqrt{\sum (\hat{y}_k - Y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
signal. Throughout this section and the next, the terms error and residual
will be referencing these values.

>>>>

To test how well the particle filter would do with plenty of data, and
determine the inherent variance in the model parameters, a very long simulation
with randomly generated impulse stimuli was created. The preprocessed timeseries
and the final estimate are shown in \autoref{fig:long_converge};
the final covariance matrix of the parameters is in \autoref{tab:long_cov}.
Note that even though the BOLD response converged (\autoref{fig:long_converge},
\autoref{fig:long_converge_500}),
the parameters still have significant correlation (\autoref{tab:long_corr}).
Based on the histograms in the first 500 seconds, the parameters converged to
their final values well before the end.
Although this is only a single test, the correlation (\autoref{tab:long_corr})
of the parameters increases the probability that the parameters are
ill-defined. When the input
consists entirely of impulses, the best parameters are not one particular
set, but a joint distribution. Note that the correlation is in parameters whose priors
are completely independent. It is possible that varying the type of input could
give improved results, although informal tests did not show significant difference.
In spite of the noisy input (green line in \autoref{fig:long_converge}), the estimates
of the BOLD were actually very close to the true (noise-free) BOLD signal.

@@

Note that throughout the results, unless otherwise specified,
the term \emph{residual} will be defined as the square root of the mean squared residual,
\begin{equation}
\text{Residual} = \sqrt{\frac{1}{N}\sum (\hat{y}_k - y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the preprocessed output
sampled at time $k$. Note the subtle distinction between this and what will be
termed the \emph{error} which is defines as
\begin{equation}
\text{Error} = \sqrt{\frac{1}{N}\sum (\hat{y}_k - Y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
signal. Throughout this section and the next, the terms error and residual
will be referencing these values.
